### 2024-09-11 article_codes.qmd
Change wilfire data (now from 2016 to 2023)

# Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

@@

# Function: get_large_polygons ----
#' Function for removing small areas from a raster file. Used for calculating large forest areas (e.g., > 5 km2).  
#' @description It returns a raster with all the polygons larger than a minimal area (YES = 1, NO = 0)
#' @param .x starts object (i.e., 1 - fuel vegetation, 0 - no fuel vegetation). 
#' @param .min_area Minimum area [in km2] for considering an vegetated area as large  (Default 5 km^2)
#' @param .resolution Resolution of the final raster (default - 100 x 100 m)
get_large_polygons <- function(.x,               
                               .min_area = 5,    
                               .resolution = 100) {

@@

# get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximum distance in [m] (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

@@

# Function: get_wui_raster ----
#' Function for WUI classification
#' @description it returns a raster with WUI areas (Interface or Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
#' the buffer area is higher than a threshold (e.g., > 50%). (YES = 1, NO = 0). 
#' @param .veg_dist raster layer indicating if the a grid cell is within a 
#' distance (e.g., 2.4 km) of a large area of flammable vegetation (YES = 1, NO = 0)
#' @param .pot_wui raster layer with potential WUI grids based on building density (`wui_potential`).
#' @param .region SpatVector layer with the region area (`region_v`) 
get_wui_raster <- function(.veg_perc,
                           .veg_dist,
                           .pot_wui = wui_potential,
                           .region = region_v){

@@

# add_colours_wui 
#' Function for adding WUI colours in a SpatRaster 
#' @description Add colours to a WUI raster layer (No_WUI, Interface, Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
add_colours_wui <- function(.x) {

@@

Then, we classified as Non-burnable the following categories:

@@

The original resolution of the map is 10 meters. We aggregate it (by a factor of 10) and transform the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.

@@

Finally, we combine both maps in one (@fig-wild-vegetation-national)

@@

Original data from [@Aragoneses2022]. Al fuels are considered for WUI. The resolution of the original map is 1km and we have re-sampled it to a grid cell of 250x250 grids (same than `region_r`) for the WUI (based in the type that is most presented in the area: `mode`).

@@

We carry out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow is therefore:

@@

Therefore, we do the following steps:

@@

Zoom to counties (@fig-wui-map-counties; @tbl-summary-wui-county):

```{r}
#| label: fig-wui-map-counties
#| fig-cap: "WUI maps by county based on different fuel/vegetation maps (resolution of 250 x 250 m)"
#| fig-subcap: 
#|  - "Viken"
#|  - "Rogaland"
#|  - "Vestland"
#|  - "Trøndelag"
#|  - "Troms og Finnmark"
#|  - "Innlandet"
#|  - "Oslo"
#|  - "Vestfold og Telemark"
#|  - "Agder"
#|  - "Møre og Romsdal"
#|  - "Nordland"
#| layout-nrow: 11
#| out-width: 100%

# plot_zoom <- function(.x) {
#   
#   zoom_area <- counties |> 
#     filter(NUTS_NAME == {{ .x }})
#   
#   wui_r_zoom <- wui_r |> 
#     crop(zoom_area) |> 
#     mask(zoom_area)
#   
#   # Add colour table 
#   coltab(wui_r_zoom) <- 
#     data.frame(id = 0:2, 
#                col = c("lightgrey", "#F0E442", "#CC79A7"))
#   
#   par(mfrow = c(1, 4))
#   
#   plot(wui_r_zoom$effis, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$effis))
#   # title( {{ .x }}, line = -4, adj = 0)
#   
#   plot(wui_r_zoom$corine, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$corine))
#   
#   plot(wui_r_zoom$firEUrisk,
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$firEUrisk))
#   
#   plot(wui_r_zoom$sr16,
#        axes = F, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$sr16)) 
#   
# }
# 
# # par(mfrow = c(1, 11))
# as.list(counties$NUTS_NAME) |> 
#   walk(plot_zoom)

@@

```
```{r}
#| label: fig-pct-wui-counties

@@

Area of wild vegetation by county (@tbl-summary-table-areas): Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution.

@@

Equivalent to select any area with building as WUI.

@@

We analyse all wildfires that have affected buildings from 2016 to 2022 (i.e., 330). The data have been obtained from XX (Norway) and XXX (Sweden).

@@

wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  mutate(datetime = lubridate::as_date(datetime)) |> 
  filter(datetime >= lubridate::as_date("2016-01-01")) |> 
  st_as_sf(coords = c("ecoord", "ncoord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

@@

Intersect points with WUI maps

@@

Get if the grid to which the wildfire is located was defined as WUI in the respective maps.

>>>>

#' Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

@@

#' Function: get_large_polygons ----
#' Function for removing small areas from a raster file. Used for calculating large forest areas (e.g., > 5 km2).  
#' @description It returns a raster with all the polygons larger than a minimal area (YES = 1, NO = 0)
#' @param .x starts object (i.e., 1 - fuel vegetation, 0 - no fuel vegetation). 
#' @param .min_area Minimum area [in km2] for considering an vegetated area as large  (Default 5 km^2)
#' @param .resolution Resolution of the final raster (default - 100 x 100 m)
get_large_polygons <- function(.x,               
                               .min_area = 5,    
                               .resolution = 100) {

@@

#' get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximum distance in [m] (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

@@

#' Function: get_wui_raster ----
#' Function for WUI classification
#' @description it returns a raster with WUI areas (Interface or Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
#' the buffer area is higher than a threshold (e.g., > 50%). (YES = 1, NO = 0). 
#' @param .veg_dist raster layer indicating if the a grid cell is within a 
#' distance (e.g., 2.4 km) of a large area of flammable vegetation (YES = 1, NO = 0)
#' @param .pot_wui raster layer with potential WUI grids based on building density (`wui_potential`).
#' @param .region SpatVector layer with the region area (`region_v`) 
get_wui_raster <- function(.veg_perc,
                           .veg_dist,
                           .pot_wui = wui_potential,
                           .region = region_v){

@@

#' add_colours_wui 
#' Function for adding WUI colours in a SpatRaster 
#' @description Add colours to a WUI raster layer (No_WUI, Interface, Intermix)
#' @param .veg_perc raster layer indicating if the percentage of vegetation in
add_colours_wui <- function(.x) {

@@

Then, we classified as non-burnable the following categories:

@@

The original resolution of the map is 10 meters. We aggregated it (by a factor of 10) and transformed the coordinate system from `SWEREF99 TM (EPSG:3006)` to `ETRS89-extended / LAEA Europe (EPSG:3035)`.

@@

Finally, we combined both maps in one (@fig-wild-vegetation-national)

@@

Original data from [@Aragoneses2022] and we considered all fuels. The resolution of the original map is 1km , so we have re-sampled it to a grid cell of 250x250 metres (same than `region_r`) based in the type that is most presented in the area; i.e., `mode`.

@@

We carried out a point-based approach with grid cells of `r raster_resolution` m x `r raster_resolution` m, and a buffer radio of `r radio_buffer` m. Our workflow was therefore:

@@

Therefore, we did the following steps:

@@

Valuer by counties: @tbl-summary-wui-county):

@@

```{r}
#| label: fig-pct-wui-counties

@@

Area of wild vegetation by county (@tbl-summary-table-areas): we calculate the number vegetation cells (i.e., grid centroid is inside the polygon), and calculate the area based on the raster resolution.

@@

Equivalent to select any area with building as potential WUI.

@@

@@

wf_se <- readxl::read_xlsx("data/wui_fires_sweden_epsg3006.xlsx") |> 
  janitor::clean_names() |> 
  filter(datetime >= 2016) |> 
  st_as_sf(coords = c("ecoord", "nccord"),
           crs =  3006) |> 
  select(geometry) |> 
  st_transform(crs = 3035) |> 
  mutate(country = "SE")

@@

We have analysed all wildfires that have affected buildings from 2016 to 2023 (i.e., `r dim(wf)[1]`; `r dim(filter(wf, country == "NO"))[1]` in Norway and \``r dim(filter(wf, country == "SE"))[1]` in Sweden). The can be obtained from XX (Norway) and XXX (Sweden).

@@

Get if the grid to which the wildfire is located was defined as WUI.
